---
author:
  - name: Jörg Müller
  
execute: 
    echo: true
    warning: false
    message: false
    
abstract: >
    This is the abstract for this chapter. 
---

# Summarising respondents data

```{r}
#| echo: false

library(tidyverse)
library(gpg)

#source("util/common.R")
source("util/encrypt.R")

# at the start of the session, need to execute this first time manually to get password callback
df.geam <- read_gpg("data/df.343696.en.rdata.gpg")

```

Introduction to the main socio-demographic variables and how to best summarise these.

Recommendations and reflections on disclosure control – when sample is small. Is sample representative of organisation or not

## Sociodemographic variables

### Gender

+----------------+-----------------------------+-------------------------+
| Code           | Question                    | Responses               |
+================+=============================+=========================+
| `SDEM004`      | Are you...                  | \[1\] A man\            |
|                |                             | \[2\] Non-binary\       |
|                |                             | \[3\] A woman\          |
|                |                             | \[4\] Prefer not to say |
+----------------+-----------------------------+-------------------------+

See the following figure

```{r}
#| echo: false
#| label: fig-sdem004
#| fig-cap: "SDEM004: Gender"
#| fig-alt: "Frequency bar plot of gender variable"

cpal <- RColorBrewer::brewer.pal(4, "Set1")

df.geam |> 
    ggplot(aes(x=SDEM004, fill=SDEM004)) + 
    geom_bar(width=.8) + 
    scale_fill_manual(values=cpal, na.value=cpal[4]) +
    guides(fill="none") +
    labs(x="", y="") +
    theme_light()

```

Text continues.

### Age

The age of respondents is capture as a integer input field (`SDEM001` - "In which year were you born?").

```{r}
#| echo: false
#| label: fig-sdem001
#| fig.cap: "Frequency chart of year of birth"


df.geam |>  
    ggplot(aes(x=SDEM001)) + 
    geom_bar()+ 
    labs(title="", y="Total") +
    theme(axis.title.x = element_blank(), 
          plot.margin = unit(c(.2,1.3,.2,.2), "cm"))
```

Most statistical analysis will work with aggregated age groups. The GEAM reporting template generates two groupings: 10 year steps, or 4 age groups: "Junior \< 30 years", "Middle 31-45 years", "Senior 46-65 years", and above "\> 65 years" of age.

::: column-margin
We know from *the first fundamental theorem of calculus* that for $x$ in $[a, b]$:

$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
:::

### Ethnic mayority or minority status

The GEAM inquires about ethnic minority or mayority status in question `SDEM002`:

+------------+---------------------------------------------------------------------------------------+-------------------------+
| Code       | Question                                                                              | Responses               |
+============+=======================================================================================+=========================+
| `SDEM002`  | Do you currently perceive yourself to be part of a majority or minority ethnic group? | \[1\] Majority\         |
|            |                                                                                       | \[2\] Minority\         |
|            |                                                                                       | \[3\] Unsure\           |
|            |                                                                                       | \[4\] Prefer not to say |
+------------+---------------------------------------------------------------------------------------+-------------------------+

As this question is not mandatory, there are likely to be a larger amount of `NA` entries that need to be removed in subsequent analysis.

```{r}
#| echo: false
#| label: fig-sdem002
#| fig.cap: "Ethnic minority /majority status"

cpal <- RColorBrewer::brewer.pal(5, "Set1")

df.geam |> 
    ggplot(aes(x=SDEM002, fill=SDEM002)) + 
    geom_bar(width=.8) + 
    scale_fill_manual(values=cpal, na.value=cpal[5]) +
    guides(fill="none") +
    labs(x="", y="Total")

```
