---
author:
  - name: Jörg Müller
  
execute: 
    echo: false
    warning: false
    message: false
abstract: >
    This is the abstract for this chapter. 
---

# Summarising respondents data

```{r}

library(tidyverse)
library(gpg)

source("util/common.R")
source("util/encrypt.R")

# at the start of the session, need to execute this first time manually to get password callback
df.geam <- read_gpg("data/df.343696.en.rdata.gpg")

```

Introduction to the main socio-demographic variables and how to best summarise these. 

Recommendations and reflections on disclosure control – when sample is small. 
Is sample representative of organisation or not


## Sociodemographic variables

### Gender

+-----------+-----------------------------------------------------------------+----------------------------+
| Code      | Question                                                        | Responses                  |
+===========+=================================================================+============================+
| `SDEM004` | Gender                                                          | [1] A man\                 |
|           |                                                                 | [2] Non-binary\            |
|           |                                                                 | [3] A woman\               |
|           |                                                                 | [4] Prefer not to say      |
+-----------+-----------------------------------------------------------------+----------------------------+

See the following figure

```{r}

#| label: fig-sdem004
#| fig-cap: "SDEM004: Gender"
#| fig-alt: "Frequency bar plot of gender variable"

cpal <- RColorBrewer::brewer.pal(4, "Set1")

df.geam %>% 
    ggplot(aes(x=SDEM004, fill=SDEM004)) + 
    geom_bar(width=.8) + 
    scale_fill_manual(values=cpal, na.value=cpal[4]) +
    guides(fill="none") +
    labs(x="", y="") +
    theme_light()

```

Text continues. 

### Age 