---
author:
  - name: Jörg Müller
  
execute: 
    echo: true
    warning: false
    message: false

abstract: >
    The following chapter provides an overview of the main socio-demographic variables of the GEAM questionnaire. It briefly introduces each variable and present the most suitable way to describe and visualise the available information. 

keywords:
  - Descriptive statistics
  - Socio-demographic variables
---

# Summarising respondents {#sec-ch02-respondents}

```{r}
#| label: setup 
#| echo: false

library(tidyverse)
library(gpg)
library(flextable)

library(knitr)
library(gtsummary)

source("util/common.R")
source("util/encrypt.R")

# at the start of the session, need to execute this first time manually to get password callback
df.geam01 <-read_gpg("data/df.343696.en.rdata.gpg")

df.geam02 <-read_gpg("data/df.784191.en.rdata.gpg") 

```

The GEAM questionnaire gathers various socio-demographic variables from respondents, including gender, age, sexual orientation, ethnic minority status, disability and/or health impairments, educational level, socio-economic class, partnership status, and transgender status. This socio-demographic information forms the basis for identifying differences between specific groups of respondents concerning selected outcomes, such as experiences of discrimination or job satisfaction. Additionally, combinations of these socio-demographic variables enable the exploration of intersectional analyses of GEAM data, revealing how certain combinations of these variables identify particularly disadvantaged sub-groups.

::: column-margin
See @sec-ch06-intersectional for a more detailed example of how to perform an intersectional, outcome oriented analysis of selected GEAM variables.
:::

In addition to socio-demographic variables, the GEAM also contains information on the current job situation of respondents. Variables describe job position, salary, type of contract, scientific discipline among others. These are briefly introduced in section 2 of this chapter. In combination with socio-demographic variables, they provide the basis for detecting wage gaps or differences in precarious working conditions.

In what follows, we briefly introduce each of the available socio-demographic variables and show how to best summarise the available information. Each variable is introduced by describing briefly its answer options. Information on the specific choices of answer options for each variable is availble in @guyanDevelopmentGenderEquality2022. 

::: callout-note
For the following chapter we use two real-world datasets, marked in the code as `df.geam01` (Organisation 1) and `df.geam02` (Organisation 2) respectively. The main difference between these datasets concerns the way how some variables have been modified and adapted to the specific organisational context. However, this does not affect their interpretability for the descriptive statistics explained in this chapter.
:::


## Socio-demographic variables

As a preparatory step it makes sense to observe the number of partial submissions of the GEAM questionnaire. Partial submissions create many `NA` entries that can be filtered out before the start of the actual analysis. 

A simple way of identifying incomplete questionnaires, consists of looking at the `lastpage` meta-data field in the questionnaire. 

An alternative consists of examining the submission date variable `submitdate`. This field is set to `NA` if the respondent did not reach the final page and press the submit button of the questionnaire. 

```{r}
#| label: check-partial

df.geam01 |>
  group_by(submitdate) |>
  summarise(Total = n()) |>
  mutate(Complete = c("Yes", "No")) |>
  table_df()

df.geam <- df.geam01 |>
  filter(!is.na(submitdate))

```



### Gender

The GEAM includes a standard question about gender identity of the respondent which allows for four answer options. Compared to other recommended approaches to ask about gender especially in the context of health research [@stadlerDiversifiedInnovationsHealth2023], this is a relatively simple yet effective approach to capture the most important differences in an organisational context.

+----------------+----------------+-----------------------------------+
| Code           | Question       | Responses                         |
+================+================+===================================+
| `SDEM004`      | Are you...     | \[1\] A man\                      |
|                |                | \[2\] Non-binary\                 |
|                |                | \[3\] A woman\                    |
|                |                | \[4\] Prefer not to say\          |
|                |                | \[5\] Prefer to self-identify as: |
+----------------+----------------+-----------------------------------+

Responses regarding gender ('SEM004') are then easily summarised with a frequency table or bar chart that demonstrates that answer option '\[2\] Non-binary' or '\[4\] Prefer not to say' or '\[5\] Prefer to self-identify as:" are comparatively few.

As the simple frequency table for 'SDEM004' demonstrates, only 2 respondents have selected '\[4\] Prefer not to say' in this example dataset with answer options '\[2\] Non-binary' and '\[5\] Other' having received no responses.

::: column-margin
The R code for producing simple frequency tables is shown for the first table of this chapter. All subsequent tables follow the same code but are hidden to improve readability of the text.
:::

```{r}
#| label: tbl-sdem004-gender
#| tbl-cap: "Frequency table of SDEM004 - Gender"
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM004)

```

::: callout-note
In the most recent GEAM version (since v3.0) **"Other"** as answer-option has been replaced with "Prefer to self-identify as" to avoid othering. As some of our data has been collected with previous GEAM versions, the "Other" label might still be visible in frequency tables and illustrations.
:::

The frequency table can then be transformed into a bar chart, indicating the absolute counts of each gender.

```{r}
#| label: fig-sdem004-gender
#| fig-cap: "Bar chart of SDEM004 - Gender"
#| fig-alt: "Bar chart of SDEM004 - Gender"
#| cap-location: margin

cpal <- RColorBrewer::brewer.pal(4, "Set1")

df.geam01 |> 
    ggplot(aes(x=SDEM004, fill=SDEM004)) + 
    geom_bar(width=.8) + 
    scale_fill_manual(values=cpal, na.value=cpal[4]) +
    guides(fill="none") +
    labs(x="", y="") +
    theme_light()

```

The GEAM reporting template constructs a binary gender variable `SDEM004.bin` which can be used for simple cross-tab analysis (see Chapter 3). 

```{r}
#| label: binary-sdem004-gender

df.geam01 <- df.geam01 |>
  mutate(SDEM004.bin = case_when(
    SDEM004 == "Prefer not to say" ~ NA,
    SDEM004 == "Non-binary" ~ NA, 
    SDEM004 == "Other" ~ NA,
    .default = SDEM004
  ))

```

### Age {#sec-summary-ch02-age}

The age of respondents is by captured as an integer variable (`SDEM001` - "In which year were you born?").

```{r}
#| echo: false
#| label: fig-sdem001-year
#| fig.cap: "Frequency chart of year of birth"
#| fig-alt: "Frequency bar plot of SDEM001 - Age variable"
#| cap-location: margin

df.geam02 |>  
    filter(!is.na(SDEM001)) |>
    ggplot(aes(x=SDEM001)) + 
    geom_bar(width=0.8) + 
    labs(x="", y="") +
    theme_light()
```

Most statistical analysis will work with aggregated age groups. The GEAM reporting template generates two groupings: 10 year steps, or 4 age groups: "Junior \<= 29 years", "Middle 30-44 years", "Senior 45-64 years", and "\>= 65 years" of age.

In R, a small transformation is necessary to convert and aggregated the year of birth into binned age groups:

```{r}
#| label: birth-convert-age 

# retrieve current year 
curyear <- as.numeric(format(Sys.Date(), format="%Y"))

# calculate age of respondent for current year
df.geam02$age <- curyear - df.geam02$SDEM001

```

Now we are ready to assign respondents to the specified age groups using the newly created `age` variable with the `cut()` command:

```{r}
#| label: age-aggregate-4g

# sub-divde age into 4 age groups:
df.geam02$age_4g <- cut(df.geam02$age,
                        c(0,29,44,64,100),
                        labels=c("Junior (<=29)", 
                                 "Middle (30-44)", 
                                 "Senior (45-64)", 
                                 "+65"))

```

And display the aggregated age groups:

```{r}
#| echo: false
#| label: fig-sdem001-age-4g
#| fig.cap: "Frequency chart of aggregated age groups"
#| fig-alt: "Frequency bar plot of aggregated age variable"
#| cap-location: margin

df.geam02 |>  
    ggplot(aes(x=age_4g)) + 
    geom_bar(width=0.8) + 
    labs(x="", y="") +
    theme_light()

```

Here is the corresponding frequency table:

```{r}
#| echo: false
#| label: tbl-age-4g
#| tbl-cap: "Frequency table of aggregated age groups"
#| cap-location: margin

df.geam02 |>
  table_frq(age_4g) 
```

Other GEAM questionnaires have modified 'SDEM001' before the launch and ask for aggregate age groups from the start. This might be advisable to protect respondents anonymity. As a consequence, the aggregation into age groups as explained above is not necessary. Here is an example:

```{r}
#| echo: false
#| label: fig-sdem001-agg
#| fig.cap: "Bar chart of SDEM001 - Aggregated age groups"
#| fig-alt: "Bar chart of SDEM001 - Aggregated age groups"
#| cap-location: margin

df.geam01 |>  
    filter(!is.na(SDEM001)) |>
    ggplot(aes(x=SDEM001)) + 
    geom_bar(width=0.8) + 
    labs(x="", y="") +
    theme_light()
```

### Current marital or partnership status

Question `SDEM006` asks about marital or partnership status which becomes especially relevant for inquires with regards to care responsibilities.

+-------------+------------------------------------------------------------------+-----------------------------------------------------------------+
| Code        | Question                                                         | Responses                                                       |
+=============+==================================================================+=================================================================+
| SDEM006     | Which best describes your current marital or partnership status? | \[1\] Single and never married or never in a civil partnership\ |
|             |                                                                  | \[2\] Married or in a civil partnership\                        |
|             |                                                                  | \[3\] Cohabiting\                                               |
|             |                                                                  | \[4\] Separated\                                                |
|             |                                                                  | \[5\] Divorced\                                                 |
|             |                                                                  | \[6\] Widowed\                                                  |
|             |                                                                  | \[7\] Prefer not to say\                                        |
|             |                                                                  | \[8\] I prefer to label my current partnership as               |
+-------------+------------------------------------------------------------------+-----------------------------------------------------------------+

```{r}
#| echo: false
#| label: tbl-sdem006-partnership
#| tbl-cap: "Frequency table of SDEM006 - Martial and partnership status"
#| cap-location: margin

df.geam02 |>
  table_frq(SDEM006) 
```

```{r}
#| echo: false
#| label: fig-sdem006
#| fig.cap: "Bar chart of SDEM006 - Martial and partnership status"
#| fig-alt: "Bar chart of SDEM006 - Martial and partnership status"
#| cap-location: margin

cpal <- RColorBrewer::brewer.pal(6, "Set1")

df.geam02 |> 
  ggplot(aes(x=SDEM006, fill=SDEM006)) + 
  geom_bar(width=.8) + 
  scale_fill_manual(values=cpal, na.value=cpal[6]) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
  guides(fill="none") +
  labs(x="", y="") + 
  theme_light()

```

Other variables of interest to be explored in relation to the partnership status concerns the care responsibilities for a dependent adult (`WCWI006`) or children (`WCWI008`). Note, that `WCWI010` also asks about being a "single parent or legal guardian" for children below the age of 18. Cross-checking `SDEM006` with `WCWI010` provides a good picture of the potential impact of caring responsibilities of respondents.

### Ethnic mayority or minority status

The GEAM inquires about ethnic minority or mayority status in question `SDEM002`:

+-----------+-----------------------------------------------------------------------------------------------------+-------------------------+
| Code      | Question                                                                                            | Responses               |
+===========+=====================================================================================================+=========================+
| `SDEM002` | Do you currently perceive yourself to be part of a majority or minority ethnic or racialized group? | \[1\] Majority\         |
|           |                                                                                                     | \[2\] Minority\         |
|           |                                                                                                     | \[3\] Unsure\           |
|           |                                                                                                     | \[4\] Prefer not to say |
+-----------+-----------------------------------------------------------------------------------------------------+-------------------------+

As this question is not mandatory, there are likely to be a larger amount of `NA` entries that need to be removed in subsequent analysis.

```{r}
#| echo: false
#| label: tbl-ethnic-sdem002
#| tbl-cap: "Frequency table of SDEM002 - Ethnic minority / majority status"
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM002) 
```

```{r}
#| echo: false
#| label: fig-ethnic-sdem002
#| fig.cap: "Bar chart of SDEM002 - Ethnic minority / majority status"
#| fig-alt: "Bar chart of SDEM002 - Ethnic minority / majority status"
#| cap-location: margin

cpal <- RColorBrewer::brewer.pal(5, "Set1")

df.geam01 |> 
  ggplot(aes(x=SDEM002, fill=SDEM002)) + 
  geom_bar(width=.8) + 
  scale_fill_manual(values=cpal, na.value=cpal[5]) +
  guides(fill="none") +
  labs(x="", y="") + 
  theme_light()

```

### Country of birth and main citizenship

Variable `SDEM012` asks about the **country of birth** of respondents. Again, depending on the privacy concerns when preparing the survey, countries can be aggregated into broader geographic regions in order to protect the anonymity of participants. This is the case in organisation 1 as shown in the following example.

```{r}
#| echo: false
#| label: tbl-country-birth
#| tbl-cap: "Frequency table of SDEM012 - Country of birth aggregated into geographic regions. Note the descending order of frequencies produced by sort.frq='desc'"
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM012) 
```

Variable `SDEM013` inquires about the **main citizenship** of respondents. Combining both variables `SDEM012` and `SDEM013` provides insights into possible migration background of respondents, given the specific country context in which the survey is implemented.

### Trans history

Question `SDEM005` inquires is a respondent "is trans or has a trans history".

+---------------+----------------------------------------------+-------------------------+
| Code          | Question                                     | Responses               |
+===============+==============================================+=========================+
| SDEM005       | Are you trans or do you have a trans history | \[1\] No\               |
|               |                                              | \[2\] Yes\              |
|               |                                              | \[3\] Prefer not to say |
+---------------+----------------------------------------------+-------------------------+

```{r}
#| echo: false
#| label: tbl-sdem005-transgender
#| tbl-cap: "Frequency table of SDEM005 - Trans history"
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM005) 
```

### Sexual orientation

Variable `SDEM007` asks respondents about their sexual orientation - which constitutes another important dimension of social discrimination.

+-------------+-----------------------------------------------+---------------------------------------------------+
| Code        | Question                                      | Responses                                         |
+=============+===============================================+===================================================+
| SDEM007     | Which best describes your sexual orientation? | \[1\] Bisexual\                                   |
|             |                                               | \[2\] Gay / lesbian\                              |
|             |                                               | \[3\] Heterosexual / straight\                    |
|             |                                               | \[4\] Prefer not to say\                          |
|             |                                               | \[5\] Another sexual orientation, please specify: |
+-------------+-----------------------------------------------+---------------------------------------------------+

```{r}
#| echo: false
#| label: tbl-sexorientation
#| tbl-cap: "Frequency table of SDEM007 - Sexual orientation. Note that answer option 'Other' has been replaced by the latest GEAM version 3 to 'Another sexual orientation'" 
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM007) 
```

Questions that allow respondents to indicate their preferred answer category are marked in the data matrix with the question code (here `SDEM007`) and the suffix `.other.` attached. As the @tbl-sexorientation indicates, two respondents indicate an additional option, which can be quickly displayed with the following code:

```{r}
#| label: code-other-sdem007
#| tbl-cap: "Extracting 'other' response options for SDEM007 - Sexual orienation"

# Extract text entries from .other. input field
df.geam01 |>
  filter(SDEM007 == "Other") |>
  select(id, SDEM007, SDEM007.other.) 
```

::: callout-warning
When analysing text input fields one should carefully check the provided information before running any type of analysis as it might contain sensitive information. A respondent might have provided a plain name which should be XXXed-out before proceeding.
:::

### Disability and/or health impairments

Variable `SDEM009` captures disability and health impairments. This is a fairly generic question that does not inquire about specific medical conditions or impairments. Asking about specific conditions makes sense in case a corresponding diversity or equality policy can address these diverse conditions with specific measures and interventions.

+-------------+-------------------------------------------------------------------------+-------------------------+
| Code        | Question                                                                | Responses               |
+=============+=========================================================================+=========================+
| SDEM009     | Do you have any disability, impairments or long term health conditions? | \[1\] No\               |
|             |                                                                         | \[2\] Yes\              |
|             |                                                                         | \[3\] Prefer not to say |
+-------------+-------------------------------------------------------------------------+-------------------------+

```{r}
#| echo: false
#| label: tbl-disability
#| tbl-cap: "Frequency table of SDEM009 - Disability and health impairments"
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM009) 
```

As can be seen from frequency @tbl-disability, among the respondents, 15 persons indicate a disabilty or health impairment.

### Educational level

The highest educational level of the respondent is covered in question `SDEM016`. Responses follow the International Standard Classification of Education (ISCED) proposed by the UNESCO Institute for Statistics [@unescoISCED2012] which provides the basis for comparing educational levels across countries.

+-------------+-----------------------------------------------------------------+-------------------------------------------------+
| Code        | Question                                                        | Responses                                       |
+=============+=================================================================+=================================================+
| SDEM016     | What is the highest qualification level that you have obtained? | \[1\] No formal education\                      |
|             |                                                                 | \[2\] Primary school / elementary school\       |
|             |                                                                 | \[3\] Secondary school / high school\           |
|             |                                                                 | \[4\] College diploma or degree\                |
|             |                                                                 | \[5\] Technical school\                         |
|             |                                                                 | \[6\] University - Baccalaureate / Bacherlor's\ |
|             |                                                                 | \[7\] University - Master's\                    |
|             |                                                                 | \[8\] University - Doctorate\                   |
|             |                                                                 | \[9\] University - Postdoctorate\               |
|             |                                                                 | \[10\] Prefer not to say\                       |
|             |                                                                 | \[11\] Other:                                   |
+-------------+-----------------------------------------------------------------+-------------------------------------------------+

```{r}
#| echo: false
#| label: tbl-edu-level
#| tbl-cap: "Frequency table of SDEM016 - Highest educational level of respondent"
#| cap-location: margin

df.geam01 |>
  table_frq(SDEM016)
```

It is likely that some respondents mark "Other" for SDEM016. These need to be examined and re-classified according to the best fit with existing categories. Examining the `SDEM016.other.` variable we see that 4 out of the 6 respondents have provided an alternative educational category, while 2 respondents have selected the "Other" option but not filled in any alternative text.

```{r}
#| label: code-other-sdem016
#| tbl-cap: "Extracting 'other' response options for SDEM016 - Educational level"

# Extract text entries from .other. input field
df.geam01 |>
  filter(SDEM016 == "Other")|>
  select(id, SDEM016, SDEM016.other.) 
```

The answer questions then can be re-coded as desired (show example).


### Socioeconomic status {#sec-ch02-ses}

Educational attainment is among the most widely used indicator for socioeconomic status (SES). Differences in SES have important implications in terms of health- or educational outcomes and thus provide an important indicator of social inequalities [@stadlerDiversifiedInnovationsHealth2023; @APAsesstatus2007]. Current SES is measured by respondents educational level (`SDEM016`). In order to track SES of entry level positions in the organisation, including early career researchers, the educational attainment of parents and/or legal guardians should be used.

Socioeconomic status is captured by two variables `SDEM017` and `SDEM018`, one for each parent and/or legal guardian. Note that answer options are indentical to `SDEM016` and follow the UNESCO ISCED standard classification.

+------------+----------------------------------------------------------------------------------+--------------------------+
| Code       | Question                                                                         | Responses                |
+============+==================================================================================+==========================+
| SDEM017    | What is the highest qualification level obtained by your first parent/guardian?  | see responses to SDEM016 |
+------------+----------------------------------------------------------------------------------+--------------------------+
| SDEM018    | What is the highest qualification level obtained by your second parent/guardian? | see responses SDEM016    |
+------------+----------------------------------------------------------------------------------+--------------------------+

Frequency table of first parent or legal tutor of respondent. 

```{r}
#| echo: false
#| label: tbl-ses-sdem017
#| tbl-cap: "Frequency table of SDEM017 - Highest educational level of parent/legal tutor"
#| cap-location: margin


df.geam01 |>
  table_frq(SDEM017)

```

And second parent / tutor: 

```{r}
#| echo: false
#| label: tbl-ses-sdem018
#| tbl-cap: "Frequency table of SDEM018 - Highest educational level of 1st parent/legal tutor"
#| cap-location: margin
#| 

df.geam01 |>
  table_frq(SDEM018)

```

In preparing the analysis of the SES of respondents, the educational level is ususally aggregated in fewer groups. This involves combining (or selecting) the educational level of parent 1 and 2. For example, one can select the higher educational level comparing parent 1 and 2 and then aggregate this highest educational level of both parents (or legal guardians) into a single variable SES.

To do so, we need some preprocessing of `SDEM017` and `SDEM018`. The following code first assigns a new value (-99) to answer options "Prefer not to say", "Other" and "NA" to make the variables comparable at all.

```{r}
#| echo: true
#| label: code-sdem01718-comp

# create new SES variable based upon the higher value of SDEM017 vs. SDEM018
df.geam01 <- df.geam01 |>
  mutate(SDEM017.comp = if_else(SDEM017 == "Prefer not to say" | 
                                SDEM017 == "Other" | 
                                is.na(SDEM017), -99, as.numeric(SDEM017)), 
         SDEM018.comp = if_else(SDEM018 == "Prefer not to say" |
                                SDEM018 == "Other" | 
                                is.na(SDEM018), -99, as.numeric(SDEM018)), 
         SES = if_else(SDEM017.comp >= SDEM018.comp, SDEM017.comp, SDEM018.comp))


```

The newly created variable `SES` now contains the hightest educational level considering both parents. The following code shows the first 6 rows of our dataset, indicating the highest educational level in the last column `SES`. 

```{r}
#| echo: false
#| label: tbl-sdem10718-comp
#| tbl-cap: "Comparing highest educational level parent 1 and parent 2"
#| cap-location: margin

df.geam01 |>
  dplyr::select(SDEM017, SDEM017.comp, SDEM018, SDEM018.comp, SES) |>
  head() |>
  table_df()

```

Then it is easy to assign the respondents socioeconomic status to three groups, consisting of "Higher SES", "Medium SES", or "Lower SES", for example. 

```{r}
#| echo: true
#| label: code-ses-3g

df.geam01 <- df.geam01 |>
  mutate(SES_3g = case_when(
    SES >0 & SES <=3 ~ "Lower SES", 
    SES >3 & SES <=5 ~ "Medium SES",
    SES >5 & SES <=9 ~ "Higher SES",
    .default = NA
  ))

```

A corresponding frequency table or bar chart is then easily created. As shown in @fig-ses-3g, most respondents have a high socio-economic background given the science and research context of the organisation. Subsequent analysis could further explore how the SES differs between the job positions captured in variable `WCJC001` (see next section).

```{r}
#| echo: false
#| label: fig-ses-3g
#| fig.cap: "Bar chart of ses_3g - Socioeconomic (parents) status"
#| fig-alt: "Bar chart of ses_3g - Socioeconomic (parents) status"
#| cap-location: margin

cpal <- RColorBrewer::brewer.pal(3, "Set1")

df.geam01 |> 
  ggplot(aes(x=SES_3g, fill=SES_3g)) + 
  geom_bar(width=.8) + 
  scale_fill_manual(values=cpal, na.value=cpal[6]) +
  guides(fill="none") +
  labs(x="", y="") + 
  theme_light()

```



## Working conditions

Respondents can also be characterised by their working conditions, including job position (`WCJC001`), teaching duties (`WCJC023`), leadership responsibilities (`WCJC027`), type of contracts (`WCJC010` and `WCJC011`), salary (`WCJC005`) and complementary bonus bonus (`WCJC005a`). Combined with the main socio-demographic variables described in the previous section, the crossing of working condition variables with the main dimensions of social discrimination such as gender, age, socioeconomic status, sexual orientation, health, and ethnic minority status provides the basic groupings to be exploreding in an initial equality audit.

### Job position

+-------------+-----------------------------------------------------------------+------------------------------+
| Code        | Question                                                        | Responses                    |
+=============+=================================================================+==============================+
| WCJC001     | What is your current position in the organization you work for? | \[1\] Academic / researcher\ |
|             |                                                                 | \[2\] Technician\            |
|             |                                                                 | \[3\] Administrative staff   |
+-------------+-----------------------------------------------------------------+------------------------------+

The categorization of staff categories follows Frascati Manual classification of R&D personnel [@oecdFrascatiManual2015 page 161ff]. 


```{r}
#| echo: false
#| label: tbl-wcjc001
#| tbl-cap: "Frequency table of WCJC001 - Job position"
#| cap-location: margin

df.geam01 |>
  table_frq(WCJC001) 
```

GEAM version 3 has follow-up questions inquiring about more fine-grade categories within each of these job positions (`WCJC024` - for Academics/researchers, `WCJC025` - for Technicians, and `WCJC0026` - for Administrative staff).

One would expect that women are overrepresented among administrative personnel while the gender represenattion should be more equal at the academic level (depending on age) and for technicians. In the following code, we also remove all `NA`s from both the gender as well as job position variable:

```{r}
#| echo: true
#| label: fig-cross-gender-by-job
#| fig.cap: "Gender by job position bar chart"
#| cap-location: margin
#| code-fold: true
#| code-summary: "Show the code"

cpal <- RColorBrewer::brewer.pal(3, "Set1")

df.geam01 |>
    filter(!is.na(WCJC001) & !is.na(SDEM004.bin)) |>
    ggplot(aes(x=WCJC001, fill=SDEM004.bin)) + 
    geom_bar(width=.8, 
             position=position_dodge()) +
    geom_text(stat='count', 
              aes(label=after_stat(count)), 
              position=position_dodge(width=.8), 
              vjust=-.6) +
    scale_fill_manual(values=cpal) +
    ylim(0,100) +
    labs(x="Job position", y="", fill="Gender") +
    theme_light()

```

Women are over-represented among administrative staff, as can be expected. Vertical segregation should be visible when examining academic positions more closely. 


### Teaching duties

A simple question asks about teaching duties of resondents (variable `WCJC023`).

::: callout-warning
Available datasets do not contain this variable!
:::


### Leadership position

Question `WCJC027` asks about the number of people that report to the respondent. This indicates leadership responsibilities.

+-------------+-----------------------------------------------------------+--------------------+
| Code        | Question                                                  | Response           |
+=============+===========================================================+====================+
| WCJC027     | How many people do directly and indirectly report to you? | \[1\] 0\           |
|             |                                                           | \[2\] 1 - 5\       |
|             |                                                           | \[3\] 6 - 10\      |
|             |                                                           | \[4\] 11 - 20\     |
|             |                                                           | \[5\] 21 - 50\     |
|             |                                                           | \[6\] More than 50 |
+-------------+-----------------------------------------------------------+--------------------+

::: callout-warning
Available datasets do not contain this variable!
:::

### Type of contract

The type of contract of respondents is covered by two separate questions. Whereas WCJC010 inquires about full-/or part-time contract, WCJC011 inquires about permanent vs. temporary contracts.

+---------------+-------------------------------------------+------------------+
| Code          | Question                                  | Responses        |
+===============+===========================================+==================+
| WCJC010       | Are you on a full- or part-time contract? | \[1\] Part-time\ |
|               |                                           | \[2\] Full-time\ |
|               |                                           | \[3\] Other:     |
+---------------+-------------------------------------------+------------------+

```{r}
#| echo: false
#| label: tbl-wcjc010
#| tbl-cap: "Frequency table of WCJC010 part-time/full-time contract"
#| cap-location: margin

df.geam01 |>
  table_frq(WCJC010) 
```

And fixed-term/temporary contract? 

+----------+---------------------------------------------------------------------+-------------------------------+
| Code     | Question                                                            | Responses                     |
+==========+=====================================================================+===============================+
| WCJC011  | Are you on a permanent/open-ended or fixed-term/temporary contract? | \[1\] Fixed-term / temporary\ |
|          |                                                                     | \[2\] Permanent / open-ended\ |
|          |                                                                     | \[3\] Other:                  |
+----------+---------------------------------------------------------------------+-------------------------------+
|          |                                                                     |                               |
+----------+---------------------------------------------------------------------+-------------------------------+

```{r}
#| echo: false
#| label: tbl-wcjc011
#| tbl-cap: "Frequency table of WCJC011 open-ended or temporay contract"
#| cap-location: margin

df.geam02 |>
  table_frq(WCJC011)
```

Type of contract, especially in terms of permanent versus fixed-term should be explored in relation to different socio-demographic variables, as it is an important indicator of precarious working conditions [@europeancommissionMORE4SupportData2021].

```{r}
#| echo: false
#| label: tbl-cross-wcjc011-sdem004-prepare
#| code-fold: true
#| code-summary: "Show preparing code for cross table"

df.sub <- df.geam02 |>
  filter(!is.na(WCJC011) & !is.na(SDEM004)) |>
  mutate(
    Gender = case_when(
      SDEM004 == "Prefer not to say" ~ NA,
      SDEM004 == "Non-binary" ~ NA, 
      SDEM004 == "Other" ~ NA,
      .default = SDEM004), 
    Contract = case_when(
      WCJC011 == "Other" ~ NA,
      .default = WCJC011) 
    ) |>
  transmute(
    Contract = fct_drop(Contract), 
    Gender = fct_drop(Gender), 
    Position = fct_drop(WCJC001)
  )

```


```{r}
#| echo: false
#| label: tbl-cross-WCJC011-SDEM004
#| tbl-cap: "Cross tab of type of contract (WCJC011) and gender (SDEM004). Cells indicate absolute numbers"
#| cap-location: margin


# from the gtsummary package. 
df.sub |>
  tbl_cross(
    row = Contract,
    col = Gender
  ) |> 
  bold_labels()

```

In this particular dataset, more women than men have a temporary contract while also more women than men have a permanent contract. 

Chapter 3 will explore in detail if these differences in type of contract between women and men are statistically significant (or not). 


```{r}
#| echo: false
#| label: tbl-cross-wcjc001-sdem004
#| tbl-cap: "Cross table of job position (WCJC001) and gender (SDEM004)"
#| cap-location: margin


df.sub |>
  tbl_cross(
    row = Position,
    col = Gender
  ) |> 
  bold_labels()

```

A final example shows potential gender differences between job position and gender. As can be seen from @tbl-cross-wcjc001-sdem004, clearly more women occupy administrative positions, while the academic research positions are 2 to 3 in favor of women. 


### Salary

```{r}
#| echo: false
#| label: tbl-wcjc005
#| tbl-cap: "Frequency table of WCJC005 - Salary"
#| cap-location: margin

df.geam01 |>
  table_frq(WCJC005)
```


## References {.unnumbered}
