---
author:
  - name: Jörg Müller

execute: 
    echo: true
    warning: false
    message: false

abstract: >
    The following chapter provides an example on how to carry out an intersectional analysis of the GEAM data, using the MAIHDA approach, Multilevel Analysis of Individual Heterogeneity and Discriminatory Accuracy. 

keywords: 
  - intersectionality
  - Latent Class Analysis
  - MAIHDA
---

# Intersectionality {#sec-ch06-intersectional}


```{r}
#| label: setup 
#| echo: false

# library(ggeffects)   # predictions and significance testing
# library(insight)     # extracting random effects variances
# library(datawizard)  # data wrangling and preparation
# library(parameters)  # model summaries
# library(performance) # model fit indices, ICC
# library(report)
#library(glmmTMB)     # multilevel modelling



library(tidyverse)
library(ggeffects)
library(lme4)
library(merTools)
#library(labelled)
library(Metrics)

library(sjmisc)
library(sjPlot)
library(gpg)         # use encryption

library(knitr)
library(kableExtra)

source("util/common.R")
source("util/encrypt.R")

# at the start of the session, need to execute this first time manually to get password callback
df.geam03 <-read_gpg("data/df.349312.en.rdata.gpg")

```


The GEAM questionnaire collects a variety of socio-demographic data that can be used for an intersectional analysis. As described in more detail in @sec-ch02-respondents, socio-demographic data include gender, age, sexual orientation, ethnic minority status, disability and/or health impairments, educational level, socio-economic class, partnership status, and transgender status. 

Following an intersectional analytical lens, we want to understand how combination of these socio-demographic categories affect specific outcomes, such as for example experiences of discrimination, harassment, job satisfaction, or other outcomes. 

Exploring intersectional effects is not limited to socio-demographic variables but can be extended to include additional variables such as for example working conditions (e.g. type of contract) or care responsibilities. 

A paragraph on intersectional theory, introduction. 


## MAIHDA 

The Multilevel Analysis of Individual Heterogeneity and Discriminatory Accuracy (MAIHDA) is a statistical approach to study how different socio-demographic and other characteristics of persons such as their neighborhood intersect and affect their wellness, health, or other outcomes of interest. The approach has been pioneered by Clare Evans and others [@evansMultilevel2018, @merloMultilevel2018] to identify intersectional effects, for example with regards to inequalities in health [@axelssonfiskChronic2018], mental health [@evansIntersectionality2019], eating disorders [@becciaEatingrelated2021] or educational outcomes [@kellerEducational2023]. Following an intersectional rationale, the MAIHDA analytical framework allows to understand better how the combination of several dimensions of discrimination can have especially aggravating consequences, for example in terms of health. 

MAIHDA uses an statistical approach that is known as multilevel modelling. Multilevel models are well established in social sciences to model the fact that individuals are not isolated atoms but pertain to nested social groupings such as a particular neighborhood, that is itself part of city which is embedded in a wider region or country. As members of these nested social structures, individuals are exposed to similar living conditions, implied for example when living in a more affluent or more deprived neighborhood, which affects the quality of life of all residents in a similar degree. 

MAIHDA uses the logic of multilevel modelling and conceives individuals (level 1) as members of certain intersectional strata (level 2). Thus a person might pertain to the strata of "women + high educational level + minority ethnic group", while another person pertains to a different stratum, such as "women + low educational level + majority ethnic group". MAIHDA allows to model how the membership in these intersectional strata affects certain outcomes of interest. This makes it particularly useful for studying inequalities in health, education, and other social outcomes.

Whereas traditional regression models can be used for exploring intersectional effects between several socio-demographic variables such as gender, age, race/ethnicity, they are limited in terms of the number of categories to combine, the sample size, model parsimony and scalability [@evansMultilevel2018]. In contrast, the MAIHDA approach partitions the variance within and between these strata, and thus can identify unique intersectional effects and provide precise estimates even for small groups.

Described as the new "gold standard" for studying intersectional inequalities [@merloMultilevel2018], several tutorials have been made available that explain the statistical background of MAIHDA  [@evansTutorial2024, @kellerEducational2023]. In what follows we adapt the R code provide by @evansTutorial2024 in order to analyse an exemplary GEAM dataset. 

::: column-margin
An additional tutorial is included in the [ggeffects package]([https://strengejacke.github.io/ggeffects/articles/practical_intersectionality.html) implements an intersectional analysis of the  [EUROFAMCARE](https://cordis.europa.eu/article/id/83657-studies-of-family-carers-of-older-people) data, a study of family cares of older people in Europe. 
::: 

::: callout-note
While the MAIHDA approach handles small N in each stratum more elegantly than standard regression models, having two few respondents within individual intersectional strata remains an issue. In the original tutorial @evansTutorial2024 generate **384 intersectional strata** combining 2 sex/gender categories x 3 race/ethnicity categories x 4 education categories x 4 income categories x 4 age categories. The MAIHDA analysis is only feasible given the dataset contains 33,000 respondents which keeps the grand majority of individual stratum above a minimum of 10 respondents. @kellerEducational2023 in comparison generate 40 intersectional strata for N=5451 students. 
::: 

GEAM employee data in higher education typically has less than N<1000 respondents, which sets practical limits on the number of intersectional strata to generate. Given practical limitations of N=632 respondents in our dataset, we limit the construction of the intersectional data to three variables:  **gender** (women/man), **disability/health** (yes/no) and **age** (three age groups), which yields 2x2x3=12 strata. Using MAIHDA approach, analyse how individuals within these strata are differently affected by experiences of microaggressions. We hypothesize that older age in combination with a disability/health condition and gender (being a women) identifies a group of employees that experience higher microaggressions. 


### Preprocess data

First, socio-demographic variables need to be preprocessed as described in @sec-ch02-respondents. In a second step, the categories of the selected socio-demographic variables are combined to form the intersectional strata. 

To start, we remove partial questionnaires. Those respondents who did not reach the end of the survey and press the submit-button have a `NA` entry in the `submitdate` and can be removed. 

::: callout-note
Partial GEAM submissions might still hold useful data, depending on which page the user abandonded the questionnaire. Depending on which combinations of variables are of interest, a more nuanced filterting should be carried out. Removing incomplete submissions does not guarantee that any of the selected variables used for further analysis does not contain missing data. 
:::

```{r}
#| label: tbl-filter-partial
#| tbl-cap: "Overivew partial and completed questionnaire submissions"


# check how many partial submissions 
df.geam03 |>
  group_by(submitdate) |>
  summarise(Total = n()) |>
  mutate(Complete = c("Yes", "No"))


# remove partial submissions
df.geam <- df.geam03 |>
  filter(!is.na(submitdate))

```

Next, **gender** will be converted to a binary variable. Given that there is only 1 "Non-binary" answer (see @tbl-sdem004-gender-raw) which is not sufficient for defining a distinct stratum, we remove "Non-binary" together with 17 "Prefer not to say" answers from the dataset for variable `SDEM004` 

```{r}
#| label: tbl-sdem004-gender-raw 
#| code-fold: false
#| tbl-cap: "Frequency table of raw SDEM004 - Gender"
#| cap-location: margin

df.geam |>
  table_frq(SDEM004)
```

The new variable is renamed from `SDEM004` to `gender`. 


```{r}
#| label: code-SDEM004-bin
#| code-fold: true
#| code-summary: Click to see preprocessing code for SDEM004

# create binary gender variable 
df.geam <- df.geam |>
  mutate(SDEM004.bin = if_else((SDEM004 == "Non-binary" | SDEM004 == "Prefer not to say"), 
                               NA_character_, SDEM004))

# re-convert to factor
df.geam$gender <- factor(df.geam$SDEM004.bin)

```


Second, **disability and health impairments** (`SDEM009`) contains 30 respondents that prefer not to answer.

```{r}
#| label: tbl-sdem009-disability-raw 
#| code-fold: false
#| tbl-cap: "Frequency table of raw SDEM009 - Disability and health impairments"
#| cap-location: margin

df.geam |>
  table_frq(SDEM009)
```

We remove those 30 respondents from the dataset before constructing the intersectional strata.  

```{r}
#| label: code-SDEM009-disability
#| code-fold: true
#| code-summary: Click to unfold preprocessing code for SDEM009

df.geam <- df.geam |>
  mutate(SDEM009.bin = if_else(SDEM009 == "Prefer not to say",
                               NA_character_, SDEM009))

# reconvert to factor
df.geam$disability <- factor(df.geam$SDEM009.bin)
```


Third, **age**. Introducing a slight variation to the procedure described in [@sec-summary-ch02-age] we have created a variable `age_3g` that aggregates the age of respondents into 3 groups: Junior (<30 years), Middle (31-45 years), and senior (> 46 years). 

```{r}
#| label: code-sdem001-age
#| code-fold: true
#| code-summary: Click to see preprocessing for age groups

# create three age groups: junior, middle, senior
df.geam <- df.geam |>
  mutate(age_3g = case_when(
    age <=30 ~ "Junior (<=30)", 
    age > 30 & age <=45 ~ "Middle (31-45)",
    age > 45  ~ "Senior (>45)",
    .default = NA
  ))

# reconvert to factor
df.geam$age_3g <- factor(df.geam$age_3g)
```

Which produces the following three age groups. The organisation has relatively few junior employees, with most employees in our sample belonging to the Senior group, with 45 or more years. 

```{r}
#| label: tbl-sdem001-age_3g
#| code-fold: false
#| tbl-cap: "Frequency table of 3 age groups"
#| cap-location: margin

df.geam |>
 table_frq(age_3g)
```


Finally, as our **outcome variable** we use items that form part of the **microaggression** scale (`BIMA001`), see @sec-annex-microaggression. For each row/respondent, we calculate the average score across all 11 microaggression items (see @sec-annex-i-microaggression for items) and store the value of each respondent into a new variable called `micagg`. Higher scores on microagression indicate that respondents have been exposed more frequently to mircoragressions than respondents with lower scores. 

```{r}
#| label: code-mean-microaggression

df.geam <- df.geam |>
  rowwise() |>
  mutate(micagg = 
           mean(
             as.numeric(
               c_across(
                 starts_with("BIMA001.SQ"))), na.rm=T))
```

This provides the following summary statistics on our three main socio-demographic variables and the outcome variable, experiences of microaggressions.   

```{r}
#| label: tbl-stats-sample
#| 

tbl1 <- df.geam |>
  count(gender)

tbl2 <- df.geam |>
  count(disability)

tbl3 <- df.geam |>
  count(age_3g)

```


### Define intersectional strata

Socio-demographic variables need to be **combined to create the intersectional strata** for the MAIHDA intersectional analysis. In total, the combination of 2 categories for gender (woman, man), 2 categories for disability (Yes, No) and three for age generates 2 x 2 x 3 = 12 strata. 

The following code creates a unique identifier for each stratum consisting of the combination of labels. `NA` entries in any of the socio-demographic variables causes the row to be removed.  

```{r}
#| label: code-create-strata
# generate strata by combining gender, disabilty and age for each row (respondent)
# any NA entry of either variable causes the row to be removed
df.geam <- df.geam |>
  filter(!is.na(gender) & !is.na(disability) & !is.na(age_3g)) |>
  mutate(strata = paste0(gender, ", ", 
                         disability, ", ", 
                         age_3g))

# convert strata into a factor variable
df.geam$strata <- factor(df.geam$strata)

# calculate size of each stratum
df.geam <- df.geam |>
  group_by(strata) |>
  mutate(n_strata = n())

# calculate percent of strata with less than 10 respondents for 12 subgroups
low <- df.geam |>
  group_by(strata) |>
  summarize(n_strata = n()) |>
  summarize(below_min = sum(n_strata < 10)/12) |>
  pull(below_min)

```

::: column-margin
For analysis with larger stratum (e.g. > 100) the combination of variable labels becomes impractical. @evansTutorial2024 recommend to geneate a numeric code where each digit represents the levels of variables. 
:::

Each respondents has been assigned the strata (or group) which correspondes to the respondents particular combination of gender, disability and age group. Thus 578 respondents (level 1) habe been assigned to 11 strata (level 2), with one stratum (A women, Yes, Junior (<=30) remaining empty. @tbl-overview-strata12 provides an overview of the number of respondents per strata. 

```{r}
#| label: tbl-overview-strata12
#| code-fold: true
#| tbl-cap: "Frequency table of 12 intersectional strata and number of cases"
#| cap-location: margin


df.geam |>
  group_by(gender, disability, age_3g) |>
  summarise(N_strata=n())

```

Given the 12 strata, 3 stratum (or groups) have fewer than 10 respondents - one case even only containing a single respondent. According @evansTutorial2024 (p.4), a lower bound of 10 respondents per stratum has been deemed acceptable, suggesting that the three strata <10 respondents should be removed (?!)

::: column-margin
For larger strata, @evansTutorial2024 recommend to calculate the percentage of strata that have less than 10 respondents. For the current example, `r low*100`% of our strata have less than 10 respondents.
:::

### MAIHDA Models


#### Model 1: simple intersectional model 

The first step of the MAIHDA approach consists of estimating a null model to partition the variance of the outcome (microaggression score) between and within intersectional strata. At this stage we operate with three different mean values: 

::: column-margin
Sample mean
: is simply the mean value across all microagression scores in our sample. In our case, it is `r round(mean(df.geam$micagg),2)`, indicating that most respondents answer either "[1] Never" or "[2] A little or rarely"

Grand mean
: calculating the mean value of the microagression score across all strata separately and taking then mean value of the stratum level means yields the *grand mean*.

Precision weighted grand mean (PWGM)
: the PWGM takes the stratum level mean values but weights these according to the size of each stratum. It therefore is a weighted average of the observed stratum means. 
:::


```{r}
#| label: code-maihda-null-model

# Fit the two-level linear regression with no covariates
model_1a <- lmer(micagg ~ (1|strata), data=df.geam)

summary(model_1a)

# store prediced microaggression scores in data frame
df.geam$m1Am <- predict(model_1a)

#broom::tidy(model_1a)

```
The null model contains only the intercept, which is interpreted as the overall average microaggression score in our dataset. To be more precise, it is the overall *precision weighted grand mean* (PWGM), which is the mean microaggression score across the mean values of each stratum, weighted by the sample size in each stratum. The PWGM (or intercept) in our example is slightly higher (1.48) than the sample mean calculated previously (1.41). 

The standard summary output of our model does not contain the Variance Partition Coefficient (VPC). 

::: column-margin
Variance Partition Coefficient (VPC)
: indicates how important the defined strata are for understanding differences in the outcome of interest, the microaggression score. As such it provides a global notion how important the intersectional strata are for understanding individual inequality. VPC values range from 0 to 1, where 0 indicates that strata are not important at all, and 1 that strata explain all differences in outcome. 
:::

The standard summary model (see above) does not indicate the VPC score for our null model, but we can use the "interclass correlation coefficient (icc)" of the `performence` package to extract this value. As indicated by @evansTutorial2024, the VPC is identical to the ICC, which indicates how similar the microaggression scores are expected to be between two randomly picked individuals from the same stratum. A high VPC/ICC value (close to 1) indicates that the microaggression scores are expected to be very similar, while a low VPC/ICC score (close to 0) indicates that their microaggression scores are very different between individuals in the same stratum. It thus indicates how strongly the intersectional strata discriminate between groups of individuals that differ with regard to the outcome. 

```{r}
#| label: code-maihda-mod1a-icc
 
performance::icc(model_1a)
```

::: column-margin
@axelssonfiskChronic2018 proposes the following cut-off values for VPC/ICC values in percent (%): 
 nonexistent (0–1) - poor (> 1 to ≤ 5) - fair (> 5 to ≤ 10) - good (> 10 to ≤ 20) - very good (> 20 to ≤ 30) - excellent (> 30)
:::

In our example, the ICC value is 0.068, i.e. around 6.8% of the microaggression variance is explained at the stratum level. According to  @axelssonfiskChronic2018 (see margin), a VPC/ICC score of 6.8% is "fair", indicating that a fair amount of the variance in the outcome is attributable to between strata level differences. For comparison, @evansTutorial2024 obtain a VPC of 9.4%, qualifying it as "a relatively large amoutn of clustering at the stratum-level" (p.9). Hence, our first model and its VPC indicate that the strata are important for explaining differences in microaggression scores between intersectional groups. In order to understand the relative importance of individual socio-demographic variables, a second model is required. 


#### Model 2: adjusted intersectional model 

The adjusted intersectional model quantifies the extend to which differences between strata level microaggression scores are due to additive versus interaction effects. 

Proportaional Change in Variance (PVC)
: measures how much between-strata variance observed in the intersectional model is explained by additive main effects vs. interaction effects. The complement of this value 1-PVC quantifies how much of the between-stratum variance remains unexplained by the additive main effects and is therefore attributable to interaction affects.

Thus, in a first step, we fit a new model (model 1b) with main additive effects, i.e. were all variables that define our strata (gender, disability, age) are used as fixed explanatory variables. 

```{r}
#| label: code-maihda-adjusted-model
#| 

model_1b <- lmer(micagg ~ gender + disability + age_3g + (1|strata), data=df.geam)

summary(model_1b)
```

The coefficients provide a first orientation on how strong the different variables affect the microaggression score. The larger (in absolute values) the coefficients, the higher the contribution of the variable to the between-stratum variance. 

We see that having a disability / health condition has the strongest effect, as it increments the microaggression score by 0.25. Gender has a relatively small effect (0.05), while age has a negative effect, reducing the microaggression score among older respondents. 

How does this new model with additive main effects change the VPC/ICC score? 

```{r}
#| label: code-maihda-mod1b-icc
 
performance::icc(model_1b)
```

::: column-margin
"The VPC now represents the proportion of the total variance that remains (after adjustment for additive effects) that is attributable to interaction effects. By attributable to interaction effects, we mean that some portion of the between-stratum variance (or inequalities) are not adequately described with consistent, additive patterns." (@evansTutorial2024, p.7)
:::

Model 1b has a much lower value of 0.006 (0.6%), indicating that the main effects explain a large portion of the between stratum variance, with some stratum inequaity being left unexplained by the additive main effects.

To improve readility, model 1a and model 1b can be combined into @tbl-maihda-mod1a-mod1b. which also incorporates the VPC/ICC values and which coefficients are significant.  

```{r}
#| label: tbl-maihda-mod1a-mod1b

#only html table. 
tab_model(model_1a, 
          model_1b, 
          show.reflvl = F, 
          p.style="stars", 
          dv.labels = c("Model 1a", "Model 1b"))
```

Age and disability appear to have a significant effect on the microaggression score. Given the above data, the stratum that comprises women only has on average a microagrression score that is 6% higher than the stratum that only contains men. Similar, strata with individuals that indicate a disability or permanent health condition have a 25% higher microaggression score than strata without a disability. And last but not least, individuals that belong to a senior age group (45 years or older) have on average a 25% lower microaggression score than the reference group (Junior 30 years or younger). 

To quantify the role of interaction effects for explaining the variance between strata, we calculate the PCV. 

```{r}
#| label: code-maihda-model2b-predict

# predict the mean outcome and confidence intervals of prediction
m1Bm <- predictInterval(model_1b, level=0.95, include.resid.var=FALSE)
# note that, unlike the stata code, this code combines the fixed and random part
# uncertainty. This saves a step later in comparison to the stata code. It also
# creates a new dataframe for these predictions, with one row per stratum.

# predict the stratum random effects and associated SEs
m1Bu <- REsim(model_1b)
```


```{r}
#| label: code-maihda-PCV

# incorporate predicted values into original data frame
df.geam <- df.geam |>
  bind_cols(m1Bm) |>
  dplyr::select(gender, 
                disability, 
                age_3g, 
                strata,
                n_strata,
                micagg, 
                m1Am,
                m1Bmfit=fit, 
                m1Bmupr=upr, 
                m1Bmlwr=lwr)


# calculate mean for all scores
df.stratum_level <- df.geam |>
  group_by(gender, disability, age_3g)|>
  summarise(across(where(is.numeric), mean), 
            .groups = "drop") |>
  mutate(rank = rank(m1Bmfit))

vc1a <-as.data.frame(VarCorr(model_1a))
vc1b <-as.data.frame(VarCorr(model_1b))


# calculate PCVs using components of these variance matrices (as percentages)
PCV1 <- ((vc1a[1,4] - vc1b[1,4]) / vc1a[1,4])*100
PCV1
```

Low PCV values imply that the main effects cannot fully explain the variation in the outcome and that the remaining between-strata variance is due to the existence of interaction effects between the social categories defining the intersectional strata. In contrast, high PCV values indicate that the main additive effects explain a large proportion of the mean-level differences between intersectional strata in the outcome. In our case the PCV indicates that > 90% of the total variance between strata is accounted for by main additive effects, and less than 10% being accounted for by interaction effects. 


### Visualising & presenting results 

A good way to present MAIHDA results illustrates the predicted microaggression scores for each intersectional strata. @tbl-predicted-stratum-means-mod1b rankes each strata according to the predicted values, placing higher microaggression scores on top and the lowest scores on the bottom. 

```{r}
#| label: tbl-predicted-stratum-means-mod1b
#| tbl-cap: "Predicted stratum means, order from highest to lowest"
#| cap-location: margin

df.stratum_level |>
  dplyr::select(-m1Am, -micagg) |>
  arrange(desc(rank))
```
Need to remove the first stratum (Man, Disability, Junior) as it only contains one respondent. Similar for other strata with N<10. 

Most affected by microaggressions are then middle and senior women with disability. 

The results of the ranked table can also be ilustrated in @fig-predicted-stratum-means-mod1b. 

```{r}
#| label: fig-predicted-stratum-means-mod1b
#| fig-cap: "Predicted stratum means - Model 1b"
#| fig-alt: "Predicted stratum means - Model 1b"
#| cap-location: margin

df.stratum_level |>
  ggplot(aes(x=rank, y=m1Bmfit)) +
  geom_point() +
  geom_pointrange(aes(ymin=m1Bmlwr, ymax=m1Bmupr)) +
  ylab("Predicted micro aggression score, Model 1B") +
  xlab("Stratum Rank") + 
  theme_bw()
```




## References
